\documentclass[11pt, a4paper]{article}
\usepackage{amsmath, amssymb, bm}
\usepackage{geometry}
\geometry{margin=2.5cm}
\setlength{\emergencystretch}{2em}

\title{Go2 Continuous Rebounce Setup and Experiment Plan}
\date{}

\begin{document}
\maketitle

\section{Current Task}

The current task is continuous rebounce control for Go2. At reset, each
environment samples a scalar target apex-height command
\begin{equation}
    h^{*} \sim \mathcal{U}(0.5, 0.8)\ \mathrm{m},
\end{equation}
and independently samples the initial drop height
\begin{equation}
    h_{\mathrm{drop}} \sim \mathcal{U}(0.5, 0.8)\ \mathrm{m}.
\end{equation}
The robot base is placed at $h_{\mathrm{drop}}$ with zero root velocity and the
default joint state. The policy must keep rebounding for the full episode. The
objective is not to jump higher forever, nor merely to recover to the initial
release height, but to make each rebound apex return close to the commanded
target height while maintaining a flat quadruped posture and staying in place.

Unlike the earlier single-rebounce setup, reaching an apex is no longer a
termination condition. The episode continues until either a failure condition is
met or the $20$ s time limit is reached. A successful rollout should therefore
show repeated cycles of
\begin{center}
    fall,\quad support interaction,\quad rebound,\quad airborne apex,\quad
    re-arm after the feet return near the support.
\end{center}

\section{Environment Variants}

Two environment IDs are used:
\begin{center}
\begin{tabular}{lll}
\hline
ID & Scene & Notes \\
\hline
\texttt{Go2-Rebounce-Flat} & rigid plane & 4096 envs, env spacing $2.5$ m \\
\texttt{Go2-Rebounce-Trampoline} & deformable trampoline & 2048 envs, env spacing $4.0$ m \\
\hline
\end{tabular}
\end{center}

The trampoline variant keeps the same rebounce MDP terms but replaces the
support with a deformable trampoline and repins the trampoline rim. Task success
does not rely on rigid-deformable contact sensors. Apex and foot-clearance gates
use only root state and kinematic foot body positions.

\section{Timing}

The simulator time step and action decimation are
\begin{equation}
    \Delta t_{\mathrm{sim}} = 0.005\ \mathrm{s},
    \qquad
    d = 4,
    \qquad
    \Delta t = d\Delta t_{\mathrm{sim}} = 0.02\ \mathrm{s}.
\end{equation}
The current training episode length is
\begin{equation}
    T_{\mathrm{episode}} = 20\ \mathrm{s}.
\end{equation}
IsaacLab multiplies every weighted reward term by $\Delta t$, and logged
\texttt{Episode\_Reward/*} values are additionally divided by
$T_{\mathrm{episode}}$.

\section{Reset and Command}

The rebounce command is
\begin{equation}
    \bm{c} = [h^{*}].
\end{equation}
In the current setup, the target rebound-apex height and the initial drop height
are sampled independently from the same range:
\begin{equation}
    h^{*} \sim \mathcal{U}(0.5,0.8),
    \qquad
    h_{\mathrm{drop}} \sim \mathcal{U}(0.5,0.8).
\end{equation}
The reset event writes $h^{*}$ into the command buffer and sets the root state
to
\begin{equation}
    \bm{p}_{\mathrm{root}} =
    \bm{o}_{\mathrm{env}} +
    \left[p_x^0,\ p_y^0,\ h_{\mathrm{drop}}+z_{\mathrm{offset}}\right],
    \qquad
    \bm{v}_{\mathrm{root}}=\bm{0},
\end{equation}
where $z_{\mathrm{offset}}=0$ in the current config. Joint positions are reset to
the default Go2 pose and joint velocities are reset to zero.

During the rollout, the target command is resampled after
\begin{equation}
    t_{\mathrm{resample}} \sim \mathcal{U}(10,20)\ \mathrm{s}.
\end{equation}
This changes only $h^{*}$ and therefore $z^{*}$; it does not teleport the robot
or change the latched initial drop height $z_0$. With a $20$ s episode, most
rollouts see zero or one mid-episode target change, which is enough to test
adaptation without turning the task into rapid command following.

The command term stores the following buffers:
\begin{center}
\begin{tabular}{lll}
\hline
Symbol & Buffer & Meaning \\
\hline
$h^{*}$ & \texttt{peak\_height} & commanded target height \\
$z^{*}$ & \texttt{target\_z} & world-frame commanded target apex height \\
$z_0$ & \texttt{drop\_z} & root height latched after reset \\
$z_{\max}$ & \texttt{peak\_z} & running root max during rebound ascent \\
$b_{\downarrow}$ & \texttt{has\_descended} & root has had negative vertical velocity \\
$b_{\uparrow}$ & \texttt{has\_rebounded} & root has had positive velocity after descent \\
\texttt{armed} & \texttt{\_apex\_armed} & next apex can be counted \\
\hline
\end{tabular}
\end{center}

\section{Observations and Actions}

The actor observation is
\begin{equation}
    \bm{o}^{\pi}_t =
    \left[
        h^{*},\,
        \bm{\omega}^{b}_t,\,
        \bm{g}^{b}_t,\,
        \tilde{\bm{q}}_t,\,
        \tilde{\dot{\bm{q}}}_t,\,
        \bm{a}_{t-1}
    \right].
\end{equation}
The critic additionally observes world root position and base linear velocity.
The policy action is a 12-DoF joint-position target action with the default Go2
joint offset. In the trampoline variant, a separate action term keeps the
trampoline rim pinned.

\section{Apex Event and Metrics}

An airborne apex is detected by a root vertical velocity sign change gated by a
geometric foot-clearance condition:
\begin{equation}
    b^{\mathrm{raw}}_t =
    b_{\downarrow,t}
    \wedge b_{\uparrow,t}
    \wedge (\dot{z}_{t-1} > 0)
    \wedge (\dot{z}_{t} \le 0),
\end{equation}
\begin{equation}
    b^{\mathrm{air}}_t =
    \mathbb{I}
    \left[
        \min_{i\in\mathcal{F}} z^{\mathrm{local}}_{i,t}
        > z_{\mathrm{surface}} + 0.08
    \right],
\end{equation}
where $\mathcal{F}$ is the set of four Go2 feet and $z_{\mathrm{surface}}=0$ in
the local support frame.

To avoid counting multiple vertical velocity zero-crossings around the same
physical apex, the apex detector is armed only once per bounce. After an
airborne apex is counted, it is disarmed. It is re-armed only when the feet
return near the support:
\begin{equation}
    \min_{i\in\mathcal{F}} z^{\mathrm{local}}_{i,t}
    \le z_{\mathrm{surface}} + 0.08.
\end{equation}
Thus an apex counted by the metric/reward/timeout logic is
\begin{equation}
    b^{\mathrm{apex}}_t =
    \texttt{armed}_t
    \wedge b^{\mathrm{raw}}_t
    \wedge b^{\mathrm{air}}_t.
\end{equation}

The logged count metrics are:
\begin{center}
\begin{tabular}{p{0.28\linewidth}p{0.30\linewidth}p{0.30\linewidth}}
\hline
Metric & Condition & Interpretation \\
\hline
\texttt{apex\_count} & $b^{\mathrm{apex}}_t$ & physical airborne apex count \\
\texttt{target\_apex\_count} &
$b^{\mathrm{apex}}_t \wedge |z_t-z^{*}|<0.25$ &
apex count close to target height \\
\texttt{error\_peak\_height} &
mean $|z_t-z^{*}|$ over counted apexes &
height tracking error \\
\hline
\end{tabular}
\end{center}

\section{Termination}

The current termination terms are:
\begin{center}
\begin{tabular}{p{0.35\linewidth}p{0.28\linewidth}p{0.27\linewidth}}
\hline
Term & Condition & Role \\
\hline
\texttt{base\_orientation} & base tilt angle $>0.6$ rad & failure \\
\texttt{non\_foot\_contact} & non-foot rigid contact $>1$ N & failure \\
\texttt{no\_airborne\_apex\_timeout} & no airborne apex for $2$ s & failure \\
\texttt{time\_out} & $20$ s reached & successful long rollout if other failures absent \\
\hline
\end{tabular}
\end{center}

\texttt{no\_airborne\_apex\_timeout} uses the same re-armed airborne-apex logic as
the metric and the apex reward. This prevents a standing policy from simply
waiting for the long time limit.

\section{Rewards}

The weighted reward is
\begin{equation}
\begin{aligned}
    R_t = \Delta t \Big(
        &-250\,r^{\mathrm{fail}}_t
        -250\,r^{\mathrm{timeout}}_t
        +5\,r^{\mathrm{progress}}_t
        +50\,r^{\mathrm{apex}}_t \\
        &-2\,r^{\mathrm{orient}}_t
        -0.5\,r^{\mathrm{inplace}}_t
        -10^{-2}\,r^{\mathrm{action}}_t
        -0.05\,r^{\mathrm{joint}}_t
        -10\,r^{\mathrm{limit}}_t
    \Big).
\end{aligned}
\end{equation}

The failure pulse is
\begin{equation}
    r^{\mathrm{fail}}_t =
    \mathbb{I}[\texttt{base\_orientation}]
    + \mathbb{I}[\texttt{non\_foot\_contact}]
    + \mathbb{I}[\texttt{no\_airborne\_apex\_timeout}].
\end{equation}
The timeout penalty is conditional:
\begin{equation}
    r^{\mathrm{timeout}}_t =
    \mathbb{I}[\texttt{time\_out}]
    \mathbb{I}[\texttt{apex\_count}<1].
\end{equation}
Thus a policy that reaches the normal time limit after producing apexes is not
penalized by \texttt{failed\_timeout}.

Let
\begin{equation}
    e^z_t = |z^{w}_{\mathrm{root},t} - z^{*}|.
\end{equation}
The flat-body gate is
\begin{equation}
    g^{\mathrm{flat}}_t =
    \exp\left(
        -\frac{\left\lVert \bm{g}^{b}_{t,xy}\right\rVert^2}{0.35^2}
    \right),
\end{equation}
and the soft foot-clearance gate is
\begin{equation}
    g^{\mathrm{feet}}_t =
    \sigma\left(
        \frac{
            \min_{i\in\mathcal{F}} z^{\mathrm{local}}_{i,t}
            -(z_{\mathrm{surface}}+0.08)
        }{0.015}
    \right).
\end{equation}

The dense rebound shaping term is
\begin{equation}
    r^{\mathrm{progress}}_t =
    \mathbb{I}[b_{\uparrow,t}]
    \exp\left(-\left(\frac{e^z_t}{0.25}\right)^2\right)
    g^{\mathrm{flat}}_t
    g^{\mathrm{feet}}_t.
\end{equation}
The apex pulse reward is
\begin{equation}
    r^{\mathrm{apex}}_t =
    \mathbb{I}[b^{\mathrm{apex}}_t]
    \exp\left(-\left(\frac{e^z_t}{0.10}\right)^2\right)
    g^{\mathrm{flat}}_t
    g^{\mathrm{feet}}_t.
\end{equation}
The remaining penalties are standard orientation, action-rate, joint-deviation,
and joint-position-limit regularizers. The in-place term penalizes drift from
the reset xy position and reset yaw:
\begin{equation}
    r^{\mathrm{inplace}}_t =
    \left\lVert
        \frac{\bm{p}_{xy,t}-\bm{p}_{xy,0}}{0.25}
    \right\rVert^2
    +
    \left(
        \frac{\mathrm{wrap}(\psi_t-\psi_0)}{0.5}
    \right)^2.
\end{equation}

\paragraph{Reward scale.}
In the current $20$ s setup, a single perfect apex pulse contributes
\[
    50 \cdot 0.02 / 20 = 0.05
\]
to the logged \texttt{Episode\_Reward/rebounce\_height}. A single failure
termination contributes approximately
\[
    -250 \cdot 0.02 / 20 = -0.25
\]
for the environments where that failure happened, before averaging over all
reset environments.

\section{Implementation Files}

\begin{center}
\begin{tabular}{ll}
\hline
File & Purpose \\
\hline
\texttt{go2\_rebounce\_env\_cfg.py} & scene, reset, reward, termination config \\
\texttt{mdp/commands.py} & rebounce command, apex metrics, re-arm state \\
\texttt{mdp/events.py} & reset from sampled drop height \\
\texttt{mdp/rewards.py} & height reward, progress reward, failure pulses \\
\texttt{mdp/terminations.py} & timeout-without-apex termination \\
\texttt{scripts/rsl\_rl/play-rebounce.py} & play script with apex height printout \\
\hline
\end{tabular}
\end{center}

\section{Experiment Roadmap}

The experiment order should separate task-definition changes from style,
efficiency, and final ablations.

\subsection{Step 0: Freeze the Current Continuous-Rebounce Baseline}

First train and save a clean baseline with the current config. Record the git
state, run ID, checkpoint, and videos. The expected healthy indicators are:
\begin{itemize}
    \item \texttt{Episode\_Termination/time\_out} close to $1$,
    \item \texttt{no\_airborne\_apex\_timeout} close to $0$,
    \item \texttt{apex\_count} close to the visually observed bounce count,
    \item \texttt{target\_apex\_count}/\texttt{apex\_count} high,
    \item \texttt{error\_peak\_height} low.
\end{itemize}

\subsection{Step 1: Decouple Initial Drop Height from Target Apex Height}

This has been implemented in the current setup. The target height and the
initial release height are no longer the same sampled scalar:
\[
    h^{*} \sim \mathcal{U}(0.5,0.8), \qquad
    h_{\mathrm{drop}} \sim \mathcal{U}(0.5,0.8).
\]
The height rewards and metrics now compare apex height against $h^{*}$ rather
than against $h_{\mathrm{drop}}$. This tests whether the policy controls
rebound height, rather than only recovering to the height from which it was
dropped.

\subsection{Step 2: Add In-Place Hopping}

This has also been implemented in the current setup with a small dense penalty
on displacement from the reset xy position and reset yaw. Suggested diagnostics
for the run are
\[
    \lVert \bm{p}_{xy,t}-\bm{p}_{xy,0}\rVert,\qquad
    \lVert \bm{v}_{xy,t}\rVert,\qquad
    |\psi_t-\psi_0|,\qquad
    |\dot{\psi}_t|.
\]
The current weight is intentionally modest so vertical rebounce quality is not
suppressed. The goal is to keep the robot rebounding in place without
sacrificing apex-height tracking.

\subsection{Step 3: Minimal Leg-Motion Regulation}

Before optimizing energy, first remove obvious high-frequency or visually
unacceptable leg motion. This should be a light regulation pass, not a strong
posture-imitation objective. The goal is to prevent wasted actuator motion and
unsafe leg swings while still allowing the robot to shape its body and legs to
use the trampoline.

Test the following regularization variants one at a time:
\begin{enumerate}
    \item increase \texttt{joint\_deviation\_l1} weight, e.g.
    $-0.05 \rightarrow -0.10$ first; avoid jumping immediately to a strong
    posture penalty because it can suppress useful trampoline-exploitation
    strategies;
    \item add a small joint-velocity penalty to reduce fast flight leg motion;
    \item reduce or reshape the foot-clearance gate to check whether the apex
    reward encourages spreading or lifting the feet near the apex.
\end{enumerate}
This step should be done before energy optimization so that the energy term is
not merely compensating for a leg-motion artifact. After each variant, check
\texttt{target\_apex\_count}, \texttt{error\_peak\_height}, in-place drift, and
videos. If height tracking or survival degrades, the regularization is too
strong.

\subsection{Step 4: Add Energy Optimization on a Fixed Trampoline}

Only after the behavior is stable and visually acceptable, add energy terms on
the fixed trampoline. Keeping trampoline parameters fixed during this step
isolates the question: can the policy maintain commanded apex height with less
active work by using the trampoline's passive energy storage?

Two mechanical-work penalties should be tested:
\[
    r^{\mathrm{pos\_work}}_t =
    \sum_j \max(\tau_{j,t}\dot{q}_{j,t}, 0),
    \qquad
    r^{\mathrm{abs\_work}}_t =
    \sum_j |\tau_{j,t}\dot{q}_{j,t}|.
\]
The positive-work term penalizes active mechanical energy injected by the
motors:
\[
    \tau_{j,t}\dot{q}_{j,t} > 0.
\]
It is the preferred first test because it discourages the robot from actively
``kicking'' to create apex height while still allowing negative work for
landing stabilization. The absolute-work term penalizes both positive and
negative mechanical work, so it also discourages active braking:
\[
    r^{\mathrm{abs\_work}}_t
    =
    r^{\mathrm{pos\_work}}_t
    +
    \sum_j \max(-\tau_{j,t}\dot{q}_{j,t},0).
\]
It should be tested after the positive-work variant, or with a much smaller
weight, because too much absolute-work penalty can reduce useful damping and
postural stabilization during trampoline compression.

A torque-squared term,
\[
    \sum_j \tau_{j,t}^{2},
\]
is also useful as a regularizer, but it mainly discourages large torques rather
than directly measuring active mechanical work.

Use a small weight or a curriculum that turns on the energy penalty only after
the policy already achieves stable rebounce. Keep the height and survival terms
dominant. If energy decreases while \texttt{target\_apex\_count},
\texttt{error\_peak\_height}, or \texttt{time\_out} degrades, the energy weight
is too large or was introduced too early. Log positive work, negative work,
absolute work, and torque norm separately; a low positive-work policy with high
negative work may simply be absorbing trampoline energy through active braking
rather than using the trampoline efficiently.

\subsection{Step 5: Add Trampoline Parameter Randomization}

After the energy-optimized policy works on the fixed trampoline, introduce
trampoline randomization. The code path already contains a reset event for
trampoline material and mass randomization, but the current ranges are fixed:
\[
    E \sim \mathcal{U}(8.0\times 10^4, 8.0\times 10^4),
    \qquad
    m \sim \mathcal{U}(10,10).
\]
Thus the current setup behaves as a fixed-trampoline experiment.

Start with narrow ranges around the fixed value, then widen them:
\[
    E \sim \mathcal{U}(0.8E_0,1.2E_0),
    \qquad
    m \sim \mathcal{U}(0.8m_0,1.2m_0),
\]
and later test wider ranges if the behavior remains stable. Randomizing too
early can make it difficult to tell whether a behavior change came from energy
optimization, leg regulation, or trampoline dynamics. It can also encourage a
more conservative active-jumping strategy that is robust but less clearly
trampoline-efficient.

To make different trampoline dynamics visible in evaluation, run fixed
evaluation sweeps over several discrete trampoline settings instead of relying
only on random training averages. For each setting, compare height error,
apex count, time-out rate, motor positive work, torque norm, contact/stance
duration, and videos. A useful result should show that the policy adapts its
timing, leg motion, and motor work across soft and stiff trampolines while
maintaining the same commanded apex-height objective.

\subsection{TODO: Real-Robot Observation Noise Measurement}

Before finalizing sim-to-real observation corruption, measure the observation
noise on the physical robot and use the measured ranges to set the policy
observation noise. The required measurements are:
\begin{itemize}
    \item joint position noise from the robot encoder readings;
    \item joint velocity noise from the robot state estimator or finite
    differences of encoder readings;
    \item IMU angular velocity noise for the base angular-velocity observation;
    \item mocap or state-estimator base position noise;
    \item mocap or state-estimator base linear-velocity noise.
\end{itemize}

For each signal, record at least a static standing segment and a representative
rebounce or hopping segment. Estimate bias, standard deviation, outliers,
latency, and any filtering delay. The resulting noise magnitudes should be used
to update the observation corruption in \texttt{go2\_rebounce\_env\_cfg.py},
especially for joint position, joint velocity, base angular velocity, base
position, and base linear velocity.

\subsection{Step 6: Final Ablation Experiments}

Run formal ablations after the final task definition is stable. Suggested
variants are:
\begin{center}
\begin{tabular}{p{0.22\linewidth}p{0.34\linewidth}p{0.34\linewidth}}
\hline
Variant & Change & Main question \\
\hline
Full & final setup & reference \\
No apex timeout & remove \texttt{no\_airborne\_apex\_timeout} & is the anti-wait gate needed? \\
No progress reward & set \texttt{rebound\_progress}=0 & is dense shaping needed? \\
No foot gate & remove foot-clearance gates & does it prevent standing or foot exploits? \\
No in-place reward & remove $xy/yaw$ terms & does the policy drift? \\
No regulation & remove flight-leg regulation & does leg motion return? \\
No energy reward & remove energy term & efficiency versus performance \\
No trampoline DR & fix trampoline parameters & robustness versus specialization \\
\hline
\end{tabular}
\end{center}

For every variant, compare success rate, failure distribution, height error,
apex count, target-apex count, drift metrics, energy, trampoline setting, and
videos.

\section{Recommended Order}

The recommended order is:
\begin{enumerate}
    \item freeze the current setup with independent drop/target heights and
    in-place constraints;
    \item add only minimal leg-motion regulation needed to remove artifacts;
    \item add energy optimization on the fixed trampoline;
    \item add trampoline parameter randomization and evaluate discrete
    trampoline settings;
    \item measure real-robot observation noise and update observation
    corruption;
    \item run final ablations.
\end{enumerate}

This order freezes the updated task definition first, removes obvious motion
artifacts, optimizes active motor energy on a known trampoline, then tests
robustness and adaptation across trampoline dynamics before final explanatory
ablations.

\end{document}
